---
title: "XML Example"
output: html_notebook
---
In this Notebook we provide a brief example of using R to search an XML document. This example comes directly from Chapter 11 "CIA Factbook Mashup" of the text "Data Science in R: A Case Studies approach to Computational Reasoning and Problem Solving" by Deborah Nolan and Dncan Temple Lang, CRC Press. The CIA Factbook Data can be downloaded from: 
http://jmatchparser.sourceforge.net/factbook/data/factbook.xml.gz

### Excerpt from the CIA Factbook XML file
```
<?xml version="1.0" encoding="UTF-8" standalone="yes"?><factbook lastupdate="2014-06-24+02:00"><news date="2014-06-26+02:00">There are dozens of monarchies scattered about the globe, bu only one of them falls in the Pacific region (Oceania). Can you name this arhcipelagic country? Hint: It lies in western Polynesia and was formerly referred to as the Friendly Islands. Its capital is Nuku'alofa, which translates as &quot;the abode of love.&quot;</news><news date="2014-06-19+02:00">Many people don't realize it, but the International Criminal Police Organization (Interpol) has been around for more than nine decades (founded 1923). How many countries are members? Check Appendix B – International Organizations and Groups to find out.</news><news date="2014-06-12+02:00">

dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="au"/><rank number="4.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="sw"/><rank number="4.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="it"/><rank number="4.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="bo"/><rank number="3.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="sn"/><rank number="3.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="gr"/><rank number="2.00" dateEstimated="false" dateLatest="2010-12-31" dateEarliest="2010-01-01" dateText="2010" country="en"/></field><field dollars="false" unit="(deaths/1,000 live births)" rankorder="1" name="Infant mortality rate" id="f2091"><description>This entry gives the number of deaths of infants under one year old in a given year per 1,000 live births in the same year. This rate is often used as an indicator of the level of health in a country.</description><rank number="117.23" dateEstimated="true" dateLatest="2014-12-31" dateEarliest="2014-01-01" dateText="2014 est." country="af"/><rank number="104.34" dateEstimated="true" dateLatest="2014-12-31" dateEarliest="2014-01-01" dateText="2014 est." country="ml"/><rank number="100.14" dateEstimated="true" dateLatest="2014-12-31" dateEarliest="2014-01-01" dateText="2014 est." country="so"/><rank number="92.86" dateEstimated="true" dateLatest="2014-12-31" dateEarliest="2014-01-01" dateText="2014 est." country="ct"/><rank number="90.92" 
```
From trial and error, and a little bit of searching we discover that information on infant mortality is contained in a <field> node which has an id attribute value of "f2091"

The rest of the notebook shows how to go about searching for infant mortality data using basic R functions to parse and search the XML document tree
```{r}
library(XML)
#url <- "http://jmatchparser.sourceforge.net/factbook/data/factbook.xml.gz"
file <- file.path(getwd(),"factbook.xml")
factbookDoc <- xmlParse(file)
factbookRoot <- xmlRoot(factbookDoc)

xmlName(factbookRoot) # confirm file
xmlSize(factbookRoot) # Find names of child nodes
table(names(factbookRoot))
# Go hunting for relevant data in the file
sapply(factbookRoot["category"],function(node) table(names(node)))
# Examine the attributes on the category nodes
sapply(factbookRoot["category"],xmlAttrs)
```
The "people and Scoiety" category looks interesting. We examine the id attribute values of all of its field children.
```{r}
categoryNodes <- factbookRoot["category"]
class(categoryNodes)
length(categoryNodes)
str(categoryNodes)
categoryNodes[[1]][1]
categoryNodes[[1]][2]

w <- sapply(categoryNodes,xmlGetAttr,"name") == "People and Society"
# Extract value of the id attributes of this node's chileren
Ids <- sapply(categoryNodes[[which(w)]]["field"],xmlGetAttr,"id")
# Search for the attribute value f2091 discovered above
f2091Index <- which(Ids == "f2091")
f2091Index
```
The desired node is the 17th <field> node of the "People and Society" <category> node
```{r}
# Find country and attribute values on the <rank> nodes
rankNodes <- categoryNodes[[which(w)]]["field"][[f2091Index]]["rank"]
xmlSize(rankNodes)
# Retrieve the valyes of the number and country attributes
infMortNum <- sapply(rankNodes,xmlGetAttr,"number")
head(infMortNum)
infMortCtry <- sapply(rankNodes,xmlGetAttr,"country")
head(infMortCtry)
```

